---
title: "The patel2014gliohuman data user's guide"
author: "Will Townes (will.townes@gmail.com) and Stephanie C. Hicks (shicks@jimmy.harvard.edu)" 
date: "April 25, 2017"
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

The `patel2014gliohuman` package contains gene expression data on 875
RNA-Seq samples from a study investigating tumor heterogeneity
in five primary glioblastoma tumors in the study
[Patel et al. (2014)](http://science.sciencemag.org/content/344/6190/1396).
This study includes both single-cell RNA-Seq samples and
bulk RNA-Seq samples from tumors and cells lines. Metadata was obtained from
Gene Expression Omnibus (GSE57872) and raw FASTQ files were
downloaded from Sequence Read Archive. The metadata includes the sequence
identifier information provided in the header of the FASTQ files which can be used
as a surrogate for batch. All samples were processed using Kallisto
for gene expression quantification. The data provided are the
estimated counts for ENSEMBL genes for all samples, not just the
ones that were included in the authors' analysis.
There are two `SummarizedExperiment` objects:
(1) estimated counts from Kallisto and (2) TPMs from Kallisto. 

The data can be accessed as follows:
```{r loadlibs, message=FALSE, warning=FALSE}
library(SummarizedExperiment)
library(patel2014gliohuman)
```

```{r loaddata}
data(patel2014gliohuman_counts) # object name: patel_glio_2014
data(patel2014gliohuman_tpm) # object name: patel_glio_2014_tpm
```

# Get expression data 

## Get the expression data in the form of counts
```{r counts} 
patel_counts = as.data.frame(as.matrix(assay(patel_glio_2014)))
patel_counts[1:5, 1:5]
dim(patel_counts)
```

## Get the expression data in the form of TPMs
```{r tpms}
patel_tpm = as.data.frame(as.matrix(assay(patel_glio_2014_tpm)))
patel_tpm[1:5, 1:5]
dim(patel_tpm)
```

Show TPMs are normalized to 1e6
```{r} 
summary(colSums(patel_tpm))
```


# Get the pheno data
```{r}
pdata = colData(patel_glio_2014)
head(pdata)
```

# References

1. Patel et al. (2014). Single-cell RNA-seq highlights
intratumoral heterogeneity in primary glioblastoma.
*Science* **344** (6190): 1396 - 1401.
PMID: 24925914 PMCID: PMC4123637.

